%%%%%%%%%%%%%%%%%%%%
%
%	Mock exam 01 : 2024--02--23 (F)
%
%%%%%%%%%%%%%%%%%%%%



\section{Exercise \ref{sec : me01q1}}
\label{sec : me01q1}

(4 pt) Let $R$ be an integral domain. For each of the following statements, state whether it is true or false, and prove it (if true) or give a counterexample (if false).
\begin{enumerate}[label=(\alph*)]
\item\label{itm : me01q1a} The product of two units is a unit.
\item\label{itm : me01q1b} If $r \in R$ is prime, then it is irreducible.
\item\label{itm : me01q1c} If $r \in R$ is irreducible, then it is prime.
\end{enumerate}

\spaceSolution{6in}{% Begin solution.
}% End solution.



\section{Exercise \ref{sec : me01q2}}
\label{sec : me01q2}

(4 pt) Let $F$ be a field, let $t$ be an indeterminate, and let $I \idealeq F[t]$ be an ideal.
\begin{enumerate}[label=(\alph*)]
\item\label{itm : me01q2a} Characterize $I$ as completely as possible.
\item\label{itm : me01q2b} For each of the following algebraic structures, state whether there exists an ideal $I \idealeq F[t]$ such that the quotient ring $F[t] / I$ has that structure. If yes, then give an example. If no, then justify why.
\begin{quote}
(i) field
\qquad
(ii) integral domain but not field
\qquad
(iii) ring but not integral domain
\end{quote}
\end{enumerate}

\spaceSolution{6in}{% Begin solution.
}% End solution.



\section{Exercise \ref{sec : me01q3}}
\label{sec : me01q3}

(4 pt) Let $\Q$ denote the field of rational numbers; given a prime $p \in \integersPositive$, let $F_{p} \isomorphism \Z / (p)$ denote the finite field with $p$ elements; and let $t$ be an indeterminate. For each of the following polynomials, state whether it is irreducible or reducible. Justify your assertions.
\begin{enumerate}[label=(\alph*)]
\item\label{itm : me01q3a} $f_{1} = t^{4} - 2 t^{3} + t + 1 \in F_{5}[t]$
\item\label{itm : me01q3b} $f_{2} = 6 t^{4} + 11 t^{3} + 8 t^{2} - 6 t - 4 \in \Q[t]$
\item\label{itm : me01q3c} $f_{3} = t^{4} + 4 t^{3} + t + 16 \in \Q[t]$
\end{enumerate}

\spaceSolution{6in}{% Begin solution.
$f_{1}$ is reducible. No zeros in $F_{5}$, but
\begin{align*}
f_{1}
=
(t^{2} + 2) (t^{2} - 2 t - 2)
\end{align*}

$f_{2}$ is reducible: In fact,
\begin{align*}
f_{2}
=
(2 t + 1) (3 t - 2) (t^{2} + 2 t + 2)
\end{align*}

$f_{3}$ is irreducible: We compute
\begin{align*}
f_{3}(t - 1)
=
t^{4} - 6 t^{2} + 9 t + 12
\end{align*}
Note that every coefficient of $f_{3}(t - 1)$ except the leading coefficient is divisible by $3$, and the constant term of $f_{3}(t - 1)$ is not divisible by $3^{2}$. Thus we may apply the Eisenstein--Sch\"{o}nemann criterion with $p = 3$ to conclude that $f_{3}$ is irreducible in $\Z[t]$ and hence, by Gau\ss{}'s lemma, in $\Q[t]$.}% End solution.



\section{Exercise \ref{sec : me01q4}}
\label{sec : me01q4}

(4 pt) Let $F$ be a field, let $G$ be a finite group, and let $V$ be an $F$-vector space.
\begin{enumerate}[label=(\alph*)]
\item\label{itm : me01q4a} Given a representation $\rho : G \rightarrow \GL(V)$, describe how to define an $F G$-module structure on $V$ that extends the $F$-action of the $F$-vector space $V$ and that affords the representation $\rho$.
\item\label{itm : me01q4a} Given an $F G$-module $V$, describe how to recover a representation $\rho$ of $G$ on $V$.
\end{enumerate}

\spaceSolution{6in}{% Begin solution.
}% End solution.



\section{Exercise \ref{sec : me01q5}}
\label{sec : me01q5}

(4 pt) Consider three matrix representations $\rho_{j} : \Z / 4 \Z \rightarrow \GL_{2}(\C)$, $j \in \{1, 2, 3\}$, defined by% Begin footnote.
\footnote{More precisely, we should write $\rho_{j}(m + 4 \Z)$. To keep notation simple, I have written what I have written.}% End footnote.
\begin{align*}
\rho_{1}(m)
&=
\begin{pmatrix}
1	&	0	\\
0	&	-1
\end{pmatrix}%
^{m}
&
\rho_{2}(m)
&=
\begin{pmatrix}
0	&	1	\\
-1	&	0
\end{pmatrix}%
^{m}
&
\rho_{3}(m)
&=
\begin{pmatrix}
i	&	0	\\
0	&	-i
\end{pmatrix}%
^{m}
\end{align*}
(Here $i \in \C$ denotes a square root of $-1$.) State which of these matrix representations are faithful, which are irreducible, and which are equivalent. Briefly justify your assertions.

\spaceSolution{6in}{% Begin solution.
}% End solution.