\section{Expectations}

\subsection{What you can expect from me}

\begin{itemize}
\item \textbf{Punctuality.} Each class will start promptly at 09h25 and end by 10h40. Assessments will be returned as promptly as is consistent with thorough review of and helpful feedback on your work. My goal is to return
\begin{itemize}
\item homework and quizzes : within 1 class meeting
\item exams : within 3 class meetings
\end{itemize}
\item \textbf{Prepared.} I will prepare each lecture with the goal of highlighting and engaging key ideas accurately and efficiently, doing my best to put myself in your shoes.
\item \textbf{Mistakes.} I will do my best to be clear and accurate. However, at times I will be unclear and make mistakes. If you find something unclear or dubious, please (politely) alert me, preferably immediately.
\end{itemize}



\subsection{What I expect from you}

\begin{itemize}
\item \textbf{Read.} Read the assigned sections before (!) each class meeting. Actively engage the reading: write your questions, comments, and quips in the margins; work through the examples. Doing so will make class meetings (and your grade) more rewarding.
\item \textbf{Participate.} Actively participate in class meetings. This may feel awkward the first few days, before we all know one another; do it anyway. Actively participating in (rather than passively attending) a class changes things, for the better.
\item \textbf{Collaborate.} Work with your colleagues on the examples in the reading, the assigned exercises, and the review problems.
\item \textbf{Communicate.} If you find something unclear or dubious, please (politely) alert me, preferably immediately.
\item \textbf{Make mistakes.} You are allowed to make mistakes. When I was learning how to ice skate, an instructor whom I greatly respected told me that if I wasn't falling on my butt multiple times each practice, then I wasn't pushing myself hard enough. I encourage you to push yourself hard and risk falling on your mathematical butt during class meetings and when working with your colleagues. If you do this, then you will soon skate cleanly, more cleanly than you might imagine.
\end{itemize}