\section{Exercise}

(5 pt) Find an equation for the plane in $\reals^{3}$ passing through the points
\begin{align*}
&(2,-2,0),
&
&(2,0,2),
&
&(0,-2,2).
\end{align*}
\fontHint{If you have trouble determining a normal vector for this plane, think whether you can use the given points to ``create'' two nonparallel vectors that lie in the plane. Then think whether you can use these to determine a normal vector for the plane.}



\spaceSolution{6in}{% Begin solution.
Geometrically, a plane is determined by (i) a point in the plane and (ii) a normal vector to the plane. We're given a point (in fact, three points) in the plane, so we just need to determine a normal vector, and then translate the geometry into algebra.

A systematic way to form a normal vector from three (noncolinear) points $A,B,C$ is to form the vectors $\fontVectorBetweenPoints{A B},\fontVectorBetweenPoints{A C}$, then compute their cross product. Doing so, we have
\begin{align*}
\fontVectorBetweenPoints{A B}
&=
(2 - 2,0 - (-2),2 - 0)
=
(0,2,2),
&
\fontVectorBetweenPoints{A C}
&=
(0 - 2,-2 - (-2),2 - 0)
=
(-2,0,2),
\end{align*}
so a normal vector $\fontVector{n}$ to the plane determined by the three points is
\begin{align*}
\fontVector{n}
&=
\fontVectorBetweenPoints{A B} \times \fontVectorBetweenPoints{A C}
\\
&=
\det
\begin{pmatrix}
\fontVector{i}	&	\fontVector{j}	&	\fontVector{k}	\\
0			&	2			&	2			\\
-2			&	0			&	2
\end{pmatrix}
\\
&=
(-1)^{1 + 1}
\det
\begin{pmatrix}
2	&	2	\\
0	&	2
\end{pmatrix}
\fontVector{i} + (-1)^{1 + 2}
\det
\begin{pmatrix}
0	&	2	\\
-2	&	2
\end{pmatrix}
\fontVector{j} + (-1)^{1 + 3}
\det
\begin{pmatrix}
0	&	2	\\
-2	&	0
\end{pmatrix}
\fontVector{k}
\\
&=
4 \fontVector{i} - 4 \fontVector{j} + 4 \fontVector{k}
=
(4,-4,4).
\end{align*}

Because the given three points are relatively ``nice'', you may also be able to ``see'' a normal vector to the plane they define: Find a corner of a room (maybe an empty room, if you're self-conscious), view the three lines emanating from that corner as the positive $x$-, negative $y$-, and positive $z$-axes (appropriately assigned), then put your feet at the point $(2,-2,0)$ (on the floor), your left hand at the point $(2,0,2)$ (on the left wall), and your right hand at the point $(0,-2,2)$ (on the right wall). If you now hold your body in a planar ``plank'' (great ab workout!), and then think about it (great mental-math workout!), you'll ``see'' that a normal vector to your body is given by $(1,-1,1)$. Note that this vector is a scalar multiple of the normal vector $\fontVector{n}$ that we found above.

Either way, we now have a normal vector for our plane. Choosing any of the given three points, say $(2,0,2)$, we can form an equation for the plane (see the general discussion below):
\begin{align*}
0
=
\fontVector{n} \cdot (x - 2,y - 0,z - 2)
=
(4,-4,4) \cdot (x - 2,y - 0,z - 2)
=
(4 x - 8) + (-4 y) + (4 z - 8).
\end{align*}
If we wish, we can isolate the $x,y,z$ terms, move the constant terms to the other side, and divide through by a common factor of $4$ to obtain
\begin{align*}
x - y + z
=
4.
\end{align*}



\paragraph{General discussion:}

Remember that a plane is determined by (i) a point in the plane and (ii) a normal vector to the plane. In particular, for a plane in $\reals^{3}$, let $P_{0} = (x_{0},y_{0},z_{0})$ be a point in the plane, and let $\fontVector{n} = (n_{1},n_{2},n_{3})$ be a normal vector to the plane. Given any point $P = (x,y,z)$ in the plane, we can form the vector from $P_{0}$ to $P$:
\begin{align*}
\fontVectorBetweenPoints{P_{0} P}
=
(x - x_{0},y - y_{0},z - z_{0}).
\end{align*}
This vector $\fontVectorBetweenPoints{P_{0} P}$ ``lives in'' the plane. (More precisely, if we move $\fontVectorBetweenPoints{P_{0} P}$ so that its initial point lies in the plane, then its terminal point --- and hence the whole vector --- lies in the plane.) Because $\fontVector{n}$ is normal (i.e. perpendicular) to the plane, we must have
\begin{align*}
\fontVector{n} \cdot \fontVectorBetweenPoints{P_{0} P}
=
0.
\end{align*}
Writing out what this equation means in terms of components, we find
\begin{align*}
0
=
\fontVector{n} \cdot \fontVectorBetweenPoints{P_{0} P}
=
(n_{1},n_{2},n_{3}) \cdot (x - x_{0},y - y_{0},z - z_{0})
=
n_{1} (x - x_{0}) + n_{2} (y - y_{0}) + n_{3} (z - z_{0}).
\end{align*}
Note that this is a scalar equation, and it is linear in the variables $x,y,z$. If desired, we can combine the scalar terms into a single term
\begin{align*}
d
=
n_{1} x_{0} + n_{2} y_{0} + n_{3} z_{0}
\end{align*}
and write the equation of the plane as
\begin{align*}
n_{1} x + n_{2} y + n_{3} z
=
d.
\end{align*}
This is the general form of an equation of a plane in $\reals^{3}$. (Can you generalize this to the general form of an equation of a plane in $\reals^{n}$?) This gives us another way to solve the above exercise: We're given three points $(x,y,z)$ that solve this equation, and we need to find the four values $n_{1},n_{2},n_{3},d$. This gives us three equations (one for each point) in four unknowns. 
}% End solution.