\documentclass[blue]{beamer}



\beamertemplatenavigationsymbolsempty
\usetheme{}
\setbeamercovered{transparent}



\begin{document}



\title{Math 357 : Class 01}
\subtitle{What should I expect in this course?}
\date{2024--01--08 (M)}



\begin{frame}
\titlepage
\end{frame}



\begin{frame}
\frametitle{Introduction}

\begin{itemize}
    \item Course: Math 357: Abstract algebra II
    \item Instructor: Stephen Wolff
    \item TA: TBD
\end{itemize}

\vspace{0.5in}

Highlights from the syllabus (below)
\end{frame}



\begin{frame}
\frametitle{To schedule}

\begin{enumerate}
    \item Office hours
    \begin{itemize}
        \item Twice weekly, each one hour plus spillover
        \item With me
    \end{itemize}
\item Recitation
    \begin{itemize}
        \item Once weekly, one hour
        \item With TA
    \end{itemize}
\end{enumerate}

Plan of action:
\begin{itemize}
    \item Propose days and times in class on Friday
    \item Start next week
    \begin{itemize}
        \item Office hours this week by appointment
    \end{itemize}
\end{itemize}
\end{frame}



\begin{frame}
\frametitle{Textbook and communication}

Textbook:
\begin{itemize}
    \item I recommend a personal copy of Dummit \& Foote (3e)
    \item Course reserves: Fondren Library Circulation Desk
\end{itemize}

\vspace{0.25in}

My communication preferences:
\begin{enumerate}
    \item In person (before or after class, during office hours)
    \item By e-mail
\end{enumerate}
\end{frame}



\begin{frame}
\frametitle{Course objectives and grading}

Two axioms:
\begin{enumerate}
    \item Goals drive grades, not vice versa.
    \item Grades are feedback for you, not judgment by me.
\end{enumerate}

\vspace{0.25in}

My goals for you:
\begin{enumerate}
    \item Over the semester, you gain comfort and confidence contributing to your mathematical community by seeing, doing, and sharing mathematics.
    \item By the end of the semester, you understand and can apply the concepts we engage.
\end{enumerate}
\end{frame}



\begin{frame}
\frametitle{Course objectives and grading}

Grading scheme:

\begin{itemize}
    \item See syllabus
    \item Focus on final destination (cf. Axiom 2 above)
    \begin{itemize}
        \item Requiz long quizzes throughout the semester
        \item Grade on each section of final exam replaces grade on corresponding midterm if former is higher
    \end{itemize}
\end{itemize}

\vspace{0.25in}

\begin{quotation}
\noindent{}This [grade] is an imperfect measure of what I know at a particular point in time. It is not a measure of who I am or who I will be.
\end{quotation}
\end{frame}



\begin{frame}
\frametitle{Policy: two remarks}

\begin{enumerate}
    \item Final exam
    \begin{itemize}
        \item It is the policy of the Mathematics Department that no final exam may be given early to accommodate student travel plans. If you make travel plans that turn out to conflict with the scheduled exam, then it is your responsibility to either reschedule your travel plans or take a zero on the exam.
    \end{itemize}
    \item Rice Honor Code (this course)
    \begin{itemize}
        \item On homework, all resources are allowed. In particular, you are strongly encouraged to work with one another. The purpose of homework is to help you to learn, share, practice, internalize concepts, and discover questions.
        \item On quizzes and exams, no external resources are allowed, unless the instructor explicitly states otherwise. The purpose of quizzes and exams is to help you to see what you can do so far, and identify what you want to work on.
    \end{itemize}
\end{enumerate}
\end{frame}



\begin{frame}
\frametitle{La r\`{e}gle du jeu}

Rule 1. Respect others and yourself.

\vspace{0.25in}

Corollaries:

\begin{enumerate}
    \item Be patient and kind, with one another and with yourself.
    \item Be on time.
    \item Be off phone.
    \item SMTP
    \begin{itemize}
        \item If you know you know, then please wait.
        \item If you have an idea, then please share.
        \item If you have no idea, then please think, listen, and ask (why?).
    \end{itemize}
\end{enumerate}
\end{frame}



\begin{frame}
\frametitle{Questions?}

\begin{enumerate}
    \item Less what answers to give, more what questions to ask?
    \item Alternative structure: Separate the exam from the course?
    \begin{itemize}
        \item Cf. qualifying exams in many graduate programs
    \end{itemize}
    \item ``Where is the wisdom we have lost in knowledge?\\
    Where is the knowledge we have lost in information?''\\
    ---T.S. Eliot, ``Choruses from the rock'' (1934)
\end{enumerate}
\end{frame}



\begin{frame}
\frametitle{Diagnostic quiz}

Basic instructions:
\begin{enumerate}
    \item Rate your confidence ($0$--$4$) for each exercise.
    \item Work on some exercises.
    \begin{itemize}
        \item If you think an exercise is unclear, then briefly (!) write your understanding and any additional hypotheses you are making, then proceed.
    \end{itemize}
    \item Submit your quiz.
\end{enumerate}

\vspace{0.25in}

Rice Honor Pledge:
\begin{quotation}
\noindent{}On my honor, I have neither given nor received any unauthorized aid on this quiz.
\end{quotation}
\end{frame}



\end{document}