%
% SQ04 : 2024--01--24 (W)
%

\noindent{}As you join your friends for another meal, one of them hands you a paper napkin on which the following has been scribbled:
\begin{align*}
f : \integers / (3) &\rightarrow \integers / (3)	\\
f(t) &= t^{3} - t
\end{align*}

\begin{enumerate}[label=(\alph*)]
\item ``The function is zero,'' one of your friends whispers in awe. Briefly provide the most convincing justification you can for this assertion.
\end{enumerate}

\spaceSolution{1in}{% Begin solution.
``It seems reasonable,'' I begin, ``to define a function to be zero if (i) its codomain has a zero element---language we typically reserve for an additive identity---and (ii) the function maps all elements of its domain to this element. The domain of our function $f$ has three elements, let us call them $-1$, $0$, and $1$.'' (A well timed arch of my eyebrow and inclination of my head stays, for the moment, objections to this choice of coset representatives.) ``One quickly confirms that $f$ indeed maps all three elements to $0$. Therefore, I share in my friend's awful assertion, if not awe.'' While my friends are momentarily befuddled by the archaic pun, I slip myself a bite to eat.}% End solution.



\begin{enumerate}[resume,label=(\alph*)]
\item ``But the polynomial is not zero in $(\integers / (3))[t]$!'' your particularly opinionated friend declaims. Briefly provide the most convincing justification you can for this assertion.
\end{enumerate}

\spaceSolution{1in}{% Begin solution.
``It also seems reasonable,'' I intervene, in less impassioned tones, ``to define a polynomial to be zero if all of its coefficients are zero. This is not the case for our polynomial $f$, which has coefficients $\pm{}1$. By definition, $1 \neq 0$ in an integral domain; and as a field, $\integers / (3)$ is an integral domain. Therefore, I also share in my friend's louder, yet no less logical, assertion.'' In a move that would make Elizabeth Bennet smile, I append a musing about truth in many forms and many voices before gracefully turning our mealtime conversation to other topics.}% End solution.