\section{Expectations}

\subsection{What you can expect from me}

\begin{itemize}
\item Punctuality. Each class will start promptly at 09h00 and end by 09h50. Quizzes and homework will be returned the following class. Exams will be returned as promptly as the logistics of a large, multi-professor course allow (expect within two weeks).
\item Prepared. I will prepare each lecture with the goal of instilling key ideas accurately and efficiently in you, doing my best to put myself in your shoes.
\item Mistakes. I will do my best to present clearly and accurately. However, almost surely at times I will be unclear and make mistakes. Please (kindly) alert me to anything that you find to be unclear or dubious, preferably immediately.
\end{itemize}



\subsection{What I expect from you}

\begin{itemize}
\item Read. Read the assigned sections before each class meeting. Actively engage the reading: write your questions, comments, and quips in the margins; work the examples (especially the ones selected in the homework document) on your own. Doing this will make class meetings (and likely your grade) more rewarding.
\item Participate. Actively participate in class discussions. This may feel awkward the first few days, before we know one another; do it anyway. It will feel natural (and rewarding!) after a few days.
\item Collaborate. Work with your peers on the examples in the reading, the assigned exercises, and the review problems.
\item Communicate. If something is unclear or dubious, please (kindly) alert me, preferably immediately.
\item Mistakes. In my classroom, you are allowed to make mistakes. When I was learning how to ice skate, an instructor whom I greatly respected told me that if I wasn't falling on my butt multiple times each session, then I wasn't pushing myself hard enough. I encourage you to push yourself hard and risk falling on your mathematical butt during regular class meetings and when working with your colleagues. If you do this, then come exam time, you will skate cleanly.
\end{itemize}